
\section{Related work}
\label{sec:relatedwork}

This section reviews the related work of logical clocks. It goes from
piggybacking approaches to vector approaches and their following compaction
approaches. Then, it reviews semantic dependency tracking and dissemination path
approaches.

\PAR{Piggybacking approaches~\cite{birman1987reliable,hadzilacos1993fault}.}{A
  trivial way to ensure causal ordering of messages is to piggyback all causally
  related messages since the last broadcast message along with the new broadcast
  message. Even by piggybacking the identifiers of messages instead of messages
  themselves, the broadcast message size may increase quickly depending on the
  application. \CBROADCAST does not piggyback all preceding messages in
  broadcast messages. However, an accumulation of messages arises during
  buffering. As discussed in Section~\ref{subsec:complexity}, we can assume that
  links quickly become safe so the buffer size stays small, and we can easily
  set a threshold on the buffer size (see Algorithm~\ref{algo:boundingbuffer}).}

\PAR{Vector clock approaches~\cite{fidge1988timestamps,mattern1989virtual}.}{A
  vector clock is a vector of monotonically increasing counters.  It encodes the
  partial order of messages using this vector:
  $VC(m) < VC(m') \implies m \rightarrow m'$.  Before delivering a message,
  processes using vector-based broadcast check if the vector of the message is
  ready regarding their local vector. If it detects any missing preceding
  message, the process delays the delivery.   To implement this
  vector-based broadcast
  \begin{inparaenum}[(i)]
  \item each process must maintain a vector locally;
  \item each message must piggyback such vector;
  \item there is 1 counter for each process of the network.
  \end{inparaenum}
  To accurately track causality, processes cannot share their
  entry. To safely track causality, processes cannot reclaim
  entries. Hence, even with \textbf{compaction
    approaches}~\cite{singhal1992efficient}, the vectors grow linearly
  in terms of number of processes that ever broadcast a message.
  In~\cite{almeida2008interval}, the complexity is reduced to the
  actual number of processes in the network.  Still, these approaches
  do not scale, particularly in dynamic networks subject to churn and
  failures. \textbf{Probabilistic
    approaches~\cite{mostefaoui2017probabilistic}} sacrifices on
  causality tracking accuracy: messages may be delivered out of order
  under a computable boundary. The size of control information in
  messages
  depends on the desired boundary. \\
  In comparison to these vector-based approaches, our approach reduces
  the generated traffic of causal broadcast by a factor of $N$ in the
  most common context where processes have partial knowledge of the
  network membership. 
% It alleviates messages from the need to piggyback heavy
  % control information~\cite{friedman2004causal}
  Unlike vector-based approach, our broadcast cannot state if two
  messages are concurrent, accurate causal delivery is a feature
  provided by default by the propagation scheme. Once a safe FIFO link
  is established, it delivers message in the causal order with no
  delay at the speed of the FIFO link.


\noindent Explicitly tracking \textbf{semantic dependencies} allows broadcast
protocols to reduce the size of piggybacked control
information~\cite{bailis2013bolton,lloyd2011cops,mukund2014optimized}. For
instance, when Alice comments Bob's picture, everyone must receive the picture
before the comment. The broadcast message only conveys one semantic
dependency. When Alice comments multiple pictures at once, the broadcast message
conveys all dependencies.  Message overhead increases linearly with the number
of semantic dependencies. To track semantic dependencies, causal broadcast
becomes application dependent. Instead our approach remains
application-agnostic. Comments, pictures, etc. are events that relate to all
preceding events. When Alice comments Bob's picture, everyone will receive this
event before the former event and all other preceding events. Whatever the
number of preceding events, broadcast messages only convey constant size control
information.

\noindent Preserving causal order using \textbf{dissemination paths} reduces
generated traffic by keeping message overhead
constant~\cite{bravo2017saturn,friedman2004causal}. State-of-the-art does not
support dynamic systems~\cite{friedman2004causal}, or supports it using
epochs~\cite{bravo2017saturn} that confines usability to small scale systems
where failures are uncommon. In comparison, we designed \CBROADCAST to handle
large and dynamic systems. Our approach provides a lightweight and efficient
mean to reconfigure dissemination paths using local knowledge without impairing
causal order.  Saturn~\cite{bravo2017saturn} along with \CBROADCAST could ease
the former's online changes in configuration.


% \PAR{Topological approaches.}{A trivial way to ensure causal delivery of
% messages without message overhead consists in building an overlay network shaped
% as a ring: each broadcast message loops once in FIFO order (\REF).
% Unfortunately, maintaining a specific topology can prove costly in dynamic
% networks subject to churn. Our approach relies on a peer-sampling protocol
% \TODO{(Relies on R-broadcast that relies on such psp)} but we make no assumption
% on the built topology. One can choose the most suitable topology depending on
% network settings. For instance, random peer-sampling
% protocols~\cite{jelasity2007gossip} have a low upkeep and guarantee a connected
% network even in dynamic settings.}

% \TODO{Explain super-peer approaches.}

% \TODO{Inter-group broadcast is a kind of topological approach.}

% \PAR{Inter-group
%   broadcast~\cite{johnson1998scalable,johnson1999intergroup}.}{Inter-group
% broadcast allows processes to broadcast messages to members of several
% inter-connected networks. Paper~\cite{johnson1999intergroup} states that
% inter-group causal broadcast is ensured when groups that internally ensure
% causal broadcast are linked together by communication channels that ensure FIFO
% ordering of messages. One can employ these approaches to cluster the network in
% subsystems. Control information in messages depends of the size of each small
% subsystem. Our approach is an extreme case where each process is a
% subsystem.}

%% The union of paths taken by broadcast messages builds a diffusion tree, i.e.,
%% a directed acyclic graph, that may change at each broadcast.

% \paragraph{Push-pull~\cite{mercier2017optimal}.} In random graphs, pull is
% optimally fast. Push phase of push-pull improves the startup phase of message
% dissemination. We only present push but it easily can be push-pull. \TODO{Maybe
%   too linked to random graphs.}


%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../paper"
%%% End:
